Red bus, blue bus

### 2021-10-12 inst/red_bus_blue_bus.Rmd
include notes

```{r loadlib}
library("hyper2")
2+4
```

>>>>

```
a,b,c in a race; a and b have an advantage S

finishing order a>b>c:   L = (a+S)/(a+b+c+2S) * (b+S)/(b+c+ S)
finishing order a>c>b:   L = (a+S)/(a+b+c+2S) * (c  )/(a+b+2S)
finishing order c>a>b:   L = (c  )/(a+b+c+2S) * (a+S)/(a+b+2S)

a,b,c,d in teams {a,b} and {c,d}

{a+b} wins, a and c have S:       (a+b+S)/(a+b+c+d+2S)
{a+b} wins, c and d have S:       (a+b)/(a+b+c+d+2S)
{a+b} wins, a and b have S:       (a+b+2S)/(a+b+c+d+2S)
{a+b} wins, a and b and c have S:  (a+b+2S)/(a+b+c+d+3S)

Car, red bus blue bus:

C  wins:  2C/(2C+RB+BB)
RB wins:  RB/(2C+RB+BB)
BB wins:  BB/(2C+RB+BB)

[NB if red and blue are really equivalent, then RB=BB=B, say and we have

C wins:  C/(C+B)
RB wins: B/(2C+2B) = 0.5*B/(C+B)
BB wins: B/(2C+2B) = 0.5*B/(C+B)

holidays:  Florida, California, California plus a nickel:

{F,C}       :  (F,C)/(F+C)  # nickel ignored in likelihood function
{F,CN}      :  (F,C)/(F+C)  # nickel ignored in likelihood function
{FN,C}      :  (F,C)/(F+C)  # nickel ignored in likelihood function
{FN,CN}     :  (F,C)/(F+C)  # nickel ignored in likelihood function

{F,FN}      :  (F,F+N)/(2F+N)
{C,CN}      :  (C,C+N)/(2C+N)

{F,C,CN}    :  (F,C,C+N)/(F+2C+N)   # FN missing
{F,FN,C}    :  (F,F+N,C)/(2F+C+N)   # CN missing

{FN,C,CN}   :  (F,C,C+N)/(F+2C+N)   # F missing
{F,FN,CN}   :  (F,F+N,C)/(2F+C+N)   # C missing

{F,FN,C,CN} :  (F,F+N,C,C+N)/(2F+2C+2N)  ??????????????

Last line (viz {F,FN,C,CN} is problematic.  Suppose that N is actually
important [as in, eg {F,FN} and N=0.999] but also N is immaterial when
comparing Florida with California.  This is a perfectly reasonable
scenario.  Then in the last line we have pretty much 50-50 probability
between {F+N,C+N}.  But this is not consistent with {F,C} which shows
that the chooser is not indifferent between Florida and California.
He might be 75% Florida and 25% California; and if this is the case we
should have {F,C} = (0.75,0.25) and {FN,CN}= (0.75,0.25) [because the
Nickel is immaterial when comparing F and C].  So what we need is for
{F,F+N,C,C+N} and N=0.999 to reduce to {F+N,C+N} which itself should
reduce to {F,C}.  And the last line above has {F,FN,C,CN} =
(0,0.5,0,0.5) which is effectively stating that he is indifferent
between Florida and California, or at least they have equal BT
strength.  This probability model is flawed because the only effect of
the nickel in {F,FN,C,CN} should be to reduce the probability of F and
C and {FN,CN} should be (F,C).

The correct probability model is to use a two-stage process as
follows:

{F,FN,FC,CN}  :  p(F)  = F/(F+C) * F/(F+N)
              :  p(FN) = F/(F+C) * N/(F+N) 
              :  p(C)  = C/(F+C) * C/(C+N) 
              :  p(CN) = C/(F+C) * N/(C+N)

Then

{F,C,CN}    :  p(F)  = F/(F+C)
            :  p(C)  = C/(F+C) * C/(C+N)
            :  p(CN) = C/(F+C) * N/(C+N)

etc.

Try this approach with red bus blue bus:

L = L(C,B,red,blue)

{C,RB}    : p(C)  = C/(C+B)
{C,RB}    : p(RB) = B/(C+B)

{C,BB}    : p(C)  = C/(C+B)
{C,BB}    : p(BB) = B/(C+B)  # (same as above)

{RB,BB}   : p(RB) =  red/(red+blue)
          : p(BB) = blue/(red+blue)

{C,RB,BB} :  p(C)  = C/(C+B) 
{C,RB,BB} :  p(RB) = B/(C+B) * red/(red+blue)
{C,RB,BB} :  p(BB) = B/(C+B) * blue/(red+blue)
```
